\documentclass[11pt]{article}
\usepackage{amsmath,amssymb}
\usepackage{geometry}
\geometry{margin=1in}

\title{The Inevitable $A/r$ Term in Coherence--Field Gravity:\
Detail, Implication, and Big Picture}
\author{Supporter Micro-Derivation}
\date{}

\begin{document}
\maketitle

% ================================================================
\section*{1. Detail: Solving the Outer Scalar Equation}
% ================================================================
In the coherence-field framework, once all baryonic mass is enclosed inside
radius $r_0$, the decoherence-weighted source $S(r)$ vanishes:
\[
S(r > r_0) = 0.
\]
The scalar equation reduces to Laplace’s equation:
\[
\nabla^2 C = 0.
\]
For spherical symmetry:
\[
\frac{1}{r^2}\frac{d}{dr}\left(r^2 \frac{dC}{dr}\right) = 0.
\]
Integrating:
\[
\frac{dC}{dr} = \frac{A_1}{r^2},
\qquad
C(r) = C_\infty + \frac{A_1}{r}.
\]

\textbf{The coherence field can only fall as $1/r$.}  
This is not a model choice — it is the unique harmonic radial solution.

% ================================================================
\section*{2. Implication: Gradient Energy Forces an $A/r$ Acceleration}
% ================================================================
The scalar contributes to the gravitational source through its gradient energy:
\[
\rho_C(r) \sim \frac{(\nabla C)^2}{C^2}
\sim \frac{A_1^2}{r^4}.
\]
In the weak-field Poisson equation,
\[
\nabla^2 \Phi = 4\pi G (\rho_b + \rho_C),
\]
the scalar component produces a potential:
\[
\Phi_C(r) \propto \ln r.
\]
Differentiation gives the induced acceleration:
\[
a_C(r) = -\frac{d\Phi_C}{dr} \propto \frac{A}{r}.
\]

Thus:
\[
a(r) = \frac{GM(r)}{r^2} + \frac{A}{r}.
\]

\textbf{This is the origin of flat rotation curves.}  
No tuning is required — the far-field coherence term \emph{must} scale this way.

% ================================================================
\section*{3. Big Picture: Why This Makes the Entire Theory Work}
% ================================================================
This one detail unlocks the entire coherence-field framework:

\begin{itemize}
\item A scale-invariant kinetic term guarantees harmonic outer behavior.
\item Harmonic outer behavior forces $C = C_\infty + A_1/r$.
\item A $1/r$ field forces a $\ln r$ potential.
\item A $\ln r$ potential forces a $1/r$ acceleration.
\item A $1/r$ acceleration is exactly what galaxies show.
\end{itemize}

This means:

\begin{itemize}
\item CFG does not fit rotation curves — it predicts them.
\item No dark matter halo profile is needed.
\item The baryonic Tully–Fisher relation follows automatically.
\item The transition radius $r_{\rm trans}$ and mass scale $M_0$ become observable.
\end{itemize}

From a single mathematical fact — the outer coherence field must be harmonic —
we obtain the entire galactic phenomenology.

This is why the coherence field is not an allowed extension of GR.

\textbf{It is the inevitable one.}

\end{document}
